---
title: "Simple document"
output: github_document:
---

```{r setup, include = FALSE}
library(tidyverse)
library(ggridges)
```

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = case_match( #create new variable "name"
      id, 
      "USW00094728" ~ "CentralPark_NY", #this weather station is in central park etc
      "USW00022534" ~ "Molokai_HI",
      "USS0023B17S" ~ "Waterhole_WA"),
    tmin = tmin / 10, #puts everything in whole degrees Celsius 
    tmax = tmax / 10) |>
  select(name, id, everything())
```

Making our first plot

```{r}
ggplot(weather_df, aes(x = tmin, y =tmax)) + 
  geom_point()
```


```{r}
weather_df |>
  #can also add a mutate |> step here then plot
  ggplot(aes( x= tmin, y = tmax))+
           geom_point()
```

```{r}
ggp_weather_scatterplot = 
  weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point()

ggp_weather_scatterplot
```

```{r}
weather_df |> 
  filter(is.na(tmax)) #shows missing tmax values
```

## Fancier 

Where you define the aesthetics matter
```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point (alpha = 0.3, size = 0.8) +#makes points a bit more transparent using alpha, makes points smaller
  geom_smooth (se = FALSE) #smooth curve to fit through scatterplot datapoints; se = standard error

```

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point (aes(color = name), alpha = 0.3, size = 0.8) +
  geom_smooth (se = FALSE)
#run this one and the one above to see difference
```

Use faceting real quick

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .3) +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name) #run to see
```

Let's make a somewhat more interesting scatterplot

```{r}
weather_df |>
  ggplot(aes(x = date, y = tmax, color = name, size = prcp)) + #set size to precipitation
  geom_point(alpha = .3) + 
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ name)
```

Learning assessment
```{r}
weather_df |> 
  filter(name == "CentralPark_NY") |> #filter
  mutate(
    tmax_fahr = tmax * (9 / 5) + 32, #convert to F
    tmin_fahr = tmin * (9 / 5) + 32 #convert to F
    ) |> #^data processing piece then plot
  ggplot(aes(x = tmin_fahr, y = tmax_fahr)) +
  geom_point(alpha = .5) + 
  geom_smooth(method = "lm", se = FALSE)
```

#small things

Just the line:
```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) + #defines aesthetics
  geom_smooth (se = FALSE) 
```

j
```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point() + 
  geom_hex() #density plot (try to install again)
```


```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) + #defines aesthetics
  geom_point(color = "blue")
  geom_smooth (se = FALSE) 
```

##univariate plots
```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, fill = name)) + 
  geom_histogram(position = "dodge") # puts bars next to each other . 
```

how do I fix this? maybe facet?

```{r}
weather_df |> 
  ggplot(aes(x = tmin)) +
  geom_histogram() + 
  facet_grid(.~name)
```

maybe a density plot?

```{r}
 ggplot(weather_df, aes(x = tmax)) + 
  geom_histogram()
```

```{r}
weather_df |> 
  ggplot(aes(x = tmin, fill = name)) +
  geom_density(alpha = .3)
```

```{r}
weather_df |> 
  ggplot(aes(x = name, y = tmin, fill = name)) +
  geom_boxplot()
```

violin plots
```{r}
weather_df |>
  ggplot(aes(x = name, y = tmin, fill = name)) + 
  geom_violin()
```

ridge plot: looks sort of like density plot
```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = name)) + 
  geom_density_ridges()
```

Learning assessment:

denisity
```{r}
ggplot(weather_df, aes(x = prcp)) + 
  geom_density(aes(fill = name), alpha = .5) 
```

Boxplot
```{r}
ggplot(weather_df, aes(y = prcp, x = name)) + 
  geom_boxplot() 
```

Ridges 1a
```{r}
ggplot(weather_df, aes(x = prcp, y = name)) + 
  geom_density_ridges(scale = .85)
```
Ridges 1b
```{r}
weather_df |> 
  filter(prcp > 0) |> 
  ggplot(aes(x = prcp, y = name)) + 
  geom_density_ridges(scale = .85)
```

```{r}
weather_df |> 
  filter(prcp > 10, prcp < 1000) |> 
  ggplot(aes(x = prcp, fill = name)) + 
  geom_density(alpha = .3)
```

## saving and embedding plots

saving plots
```{r}
ggp_weather = 
  weather_df |> 
  ggplot(aes(x = date, y = tmax, color = name))+
  geom_point()

ggsave("ggp_weather.pdf", ggp_weather, width = 8, height = 6)  #will save the most recent figure plot by default, or can specify as above with dimensions
```

embedding plots
```{r, fig.width = 12, fig.height = 12}
weather_df |> 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point()
#adding fig.width or height will change the embedded figure 
#or can use "fig.width = 6, fig.asp = 6" to set it as high as 
```

